---
title: "Covid91 vaccine study Final2020"
author: "Adesuwa Mukoro!"
date: "october 24, 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

$H_0$: The null hypothesis is that among the target population, the vaccine has no effect on whether one contracts the virus or not.

$H_A$: The alternative hypothesis is that among the target population, the vaccine has an effect on whether one contracts the virus or not.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Every question is of a cat~cat type

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

##Graphical distriptive result

```{r}
DataF=filter(FinalData,sex=="gal")
barchartGC(~infected + treatment,data=DataF)
barchartGC(~infected + treatment,data=DataF, type="percent",main="gals")

```

Among the females who contracted aids, twice as many took the placebo. Therefore, it appears that the vaccine does have an effect.

```{r}
DataG=filter(FinalData,sex=="guy")
barchartGC(~infected + treatment,data=DataG)
barchartGC(~infected + treatment,data=DataG, type="percent",main="guys")

```

Among the males who contracted covid-19, more of the guys have the vaccine. Therefore, it appears that the vaccine does have a positive effect. Taking the vaccine makes guys more likely to contract covid-19.

```{r}
DataLGBTQ=filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=DataLGBTQ)
barchartGC(~infected + treatment,data=DataLGBTQ, type="percent",main="gays")

```

Among the gays that contracted covid-19, more of them had the vaccine. The vaccine is not effective in preventing gays from contracting covid-19.

```{r}
DataU=filter(FinalData,DrugUser=="yes")
barchartGC(~infected + treatment,data=DataU)
barchartGC(~infected + treatment,data=DataU, type="percent",main="DrugUser")

```

Among the DrugUser 30% who had the virus took the vaccine and 65% who had the virus did not take the vaccine. Comparing those who did not contract the virus to the DrugUser who did,it appears there is a difference between those who did or did not take the vaccine.

##Numerical graphical summaries

```{r}
table1 <- xtabs(~infected + treatment,data=DataF)
rowPerc(table1)
colPerc(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fisher test shows the odd of getting covid-19 are 0.49 to 1. The confidence interval for the odd ratio is 0.41 to 0.58. This means that the females who got the vaccine are half as likely to contract covid-19 as females who got the placebo because the 95% confidence interval is less than 1, the result is significant.

```{r}
table1 <- xtabs(~infected + treatment,data=DataG)
rowPerc(table1)
colPerc(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Fisher test shows that the odds of getting covid-19 are 1.47 to 1. The confidence interval for the odd ratio is 1.18 to 1.85.The interval is greater than 1, so it is significant. This means guys who got the vaccine are 1.5 times more likely to contract covid-19 than those who got the placebo.

```{r}
table1 <- xtabs(~infected + treatment,data=DataLGBTQ)
rowPerc(table1)
colPerc(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Fisher test shows the odd ratio of getting covid-19 are 2.49 to 1. The 95% confidence interval for the odd ratio is 1.63 to 3.88. The interval is greater than 1, so it is significant. This means that the gays who took the vaccine,2.5 times more likely to contract covid-19 than those who took the placebo.

```{r}
table1 <- xtabs(~infected + treatment,data=DataU)
rowPerc(table1)
colPerc(table1)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fisher test shows the odds of getting covid-19 are 0.55 to 1. The 95% confidence interval for the odd ratio is 0.36 to 0.82. This interval is less than 1, so the result is significant. This means that the DrugUsers who took the vaccine are half as likely to contract covid-19 than those who took the placebo.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The vaccine appears to be useful in protecting females fron contracting covid-19

The vaccine appears to have a positive effect on the guys. Among the guys who took the vaccine, more of them contracted covid-19. It appears that the vaccine is not effective in preventing them from contracting covid-19.

Among the gays, it appears that the vaccine is not effective in preventing them from contracting the virus. 

Among the DrugUser,30% who had the virus took the vaccine and 65% who had the virus did not take the vaccine. it appears that the vaccine is effective in helping to prevent them from contracting covid-19.
